---
title: "Compare annotators"
output: 
  html_notebook: 
    number_sections: yes
---

# Setup

```{r setup}
library(credibility)
library(webannor)
library(tidyverse)
library(magrittr)
library(stringr)
```


# Import annotations

Import annotations from project directory exported by WebAnno.

```{r}
path <- file.path(getwd(), "../data-raw", "annotations", "2017-04-26")
annotations <- 
  get_annotations.WebAnno_project(path) %>% 
  ### A: could abstract this step into a function to replace annotation field names, and make the replacements to be parameters to the program, rather than hardwired
  mutate(annotation_type = stringr::str_replace(annotation_type, "Score", "sentiment"),
         annotation_type = stringr::str_replace(annotation_type, "Context", "topic"),
         annotation_type = stringr::str_replace(annotation_type, "hypothetical", "nature"))
annotations
```

Where the tagging policy is to have a default tag implied by no tag, we need to impute this to the dataframe. In case there are documents which *no* annotator has explicitly tagged, we need an external list of documents shown to annotators.

```{r}
## A: impute topics to blanks -- Not BoE
## A: get list of documents input to WebAnno so can also impute over docs lacking any annotations at all
```


Enumerate the set of annotation types that have been applied to this corpus.

```{r}
unique(annotations$annotation_type)
```



# Topics(s)

Extract topics tags.

```{r}
### A: see if can map over different annotation_types, rather than repeating this pattern for each
topics <- 
  annotations %>% 
  filter(annotation_type == "topic") %>% 
  select(doc_id, id_annotation, annotator, annotation_value)
```

Enumerate topics used (not necessarily the full set available for use).

```{r}
get_tagset_used(topics)
# topics_list <- unique(topics$annotation_value)
# topics_list
```




## Marginal distribution by annotator

Calculate within-annotator topic share.

```{r}
### Now done directly in ggplot
# topic_shares <- 
#   topics %>%
#   group_by(annotator, annotation_value) %>% 
#   summarise(ntag = n()) %>% 
#   group_by(annotator) %>% 
#   mutate(topic_share = ntag / sum(ntag))
```

Add the mean over annotators.

```{r}
# topic_shares_all <- 
#   topics %>% 
#   group_by(annotation_value) %>% 
#   summarise(ntag = n()) %>% 
#   mutate(topic_share = ntag / sum(ntag),
#          annotator = "(all)")
# 
# topic_shares %<>% 
#   bind_rows(topic_shares_all)
# 
# # order topics in descending order of frequency in (all)
# topics_ordered <- 
#   topic_shares_all %>% 
#   arrange(desc(topic_share)) %>% 
#   select(annotation_value) %>% 
#   unlist()
# 
# topic_shares

# Simpler code to order topics by overall popularity
tags_ordered_by_decreasing_popularity <- 
  topics %>% 
  group_by(annotation_value) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n)) %>% 
  select(annotation_value) %>% 
  unlist(use.names = FALSE)

```


Visualise.

```{r, fig.height=6}
# topic_shares %>%  
#   ggplot(aes(x = annotation_value)) +
#   geom_bar(aes(y = topic_share), stat = "identity") +
#   facet_grid(annotator ~ .) +
#   theme(axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5)) +
#   scale_x_discrete(name = NULL,
#                    labels = function(x) str_wrap(x, width = 15),
#                    limits = topics_ordered) +
#   scale_y_continuous(name = "Share of all topic tags applied")

# topics %>%
#   ggplot(aes(x = annotation_value)) +
#   geom_bar(aes(y = ..prop.., group = 1)) + # group = 1 needed to get proportions within annotator, rather than as proportion of margin 
#   facet_grid(annotator ~ ., margins = TRUE) +
#   theme(axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5)) +
#   scale_x_discrete(name = NULL,
#                    labels = function(x) str_wrap(x, width = 15),
#                    limits = tags_ordered_by_decreasing_popularity) +
#   scale_y_continuous(name = "Share of tags by annotator")

get_tags_ordered_by_decreasing_popularity <- function(annotations, tags = NULL) {
  # get (sorted) list of tags used in annotations
  tags_used <- 
      annotations %>% 
      group_by(annotation_value) %>% 
      summarise(n = n()) %>% 
      arrange(desc(n)) %>% 
      select(annotation_value) %>% 
      unlist(use.names = FALSE)
  # if an explicit list of possible tags is given then check for...
  if (!is.null(tags)) {
    # ... tags used in the annotations that aren't also in `tags`
    pretty_print_tags <- function(tags) paste0(tags_unexpected, "", collapse = "\n")
    tags_unexpected <- setdiff(tags_used, tags)
    if (length(tags_unexpected) > 0) warning("Tags in `annotations` that were not in `tags`: ", pretty_print_tags(tags_unexpected))
    # ... unused tags
    tags_unused <- setdiff(tags, tags_used)
    if (length(tags_unexpected) > 0) warning("Tags in `tags` that were not used in `annotations`: ", pretty_print_tags(tags_unused))
    tags <- c(tags_used, tags_unused)
  } else {
    tags <- tags_used
  } 
  tags
}

plot_marginal_tag_use_by_annotator <- 
  function(annotations, tags = NULL, tag_order = c("value", "popularity")) {
  
  tag_order <- match.arg(tag_order)
  tags_ordered <- switch(tag_order, 
                         popularity = get_tags_ordered_by_decreasing_popularity(annotations),
                         value = sort(unique(annotations$annotation_value)))  
  
  annotations %>%
    ggplot(aes(x = annotation_value)) +
    geom_bar(aes(y = ..prop.., group = 1)) + # group = 1 needed to get proportions within annotator, rather than as proportion of margin 
    facet_grid(annotator ~ ., margins = TRUE) +
    theme(axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5)) +
    scale_x_discrete(name = NULL,
                     labels = function(x) str_wrap(x, width = 15),
                     limits = tags_ordered) +
    scale_y_continuous(name = "Share of tags by annotator")
}
```

Dropping non-BoE tags...

```{r, fig.height=6}
### WARNING! Hardwired topic tag strings
non_BoE_tags <- c("Government", "Not BoE or government")
topic_shares %>%  
  filter(!(annotation_value %in% non_BoE_tags)) %>% 
  ggplot(aes(x = annotation_value)) +
  geom_bar(aes(y = topic_share), stat = "identity") +
  facet_grid(annotator ~ .) +
  theme(axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5)) +
  scale_x_discrete(name = NULL,
                   labels = function(x) str_wrap(x, width = 15),
                   limits = topics_ordered[!(topics_ordered %in% non_BoE_tags)]) +
  scale_y_continuous(name = "Share of all topic tags applied")
```


## Co-occurence of tags within annotator (multiple tagging)

Reusable plotting function.

```{r}
# Note: could make the below less specific to annotation

annotators_subset <- function(annotations, annotators = NULL) {
  # Default to all annotators
  # WARNING! Hardwired column name "annotator"
  if (!is.null(annotators)) df <- dplyr::filter(annotations, annotator %in% annotators)
  else df <- annotations
  df
}

# Take a list of annotations, by document, annotator, and annotation value (tag), and calculate agreement among a specified subset of annotators.
### Q: Should we complete missing pairs with zeros? Or provide an option to do so, with corresponding function argument?
agreement <- function(annotations, annotators = NULL) {
  df <- annotators_subset(annotations, annotators)
  # WARNING! Hardwired column name "annotation_value", "doc_id"
  widyr::pairwise_count(df, annotation_value, doc_id, diag = TRUE)
}

get_annotators <- function(annotations) unique(annotations$annotator)

### A: check that item1 corresponds to annotator1 and likewise for 2
agreement_pair <- function(annotations, annotators) {
  stopifnot(is.character(annotators) && length(annotators) == 2)
  df <- agreement(annotations, annotators)
  df$annotator1 = annotators[1]
  df$annotator2 = annotators[2]
  df
}

# Row-bind together all pairwise combinations of annotators, plus optionally margins
agreement_pairwise <- function(annotations, annotators = NULL, diag = TRUE) {
  # If an annotator is listed in the annotators argument, but doesn't appear in the annotations dataset, that annotator will be imputed to exist and implicitly have 
  df <- annotators_subset(annotations, annotators)
  # If annotators argument is left NULL then infer annotators list from annotations dataframe (so won't see annotators who exist but didn't contribute any annotations)
  if (is.null(annotators)) annotators <- get_annotators(annotations)
  annotator_pairs <- 
    if (diag) {
      temp <- gtools::permutations(length(annotators), 2, annotators, repeats.allowed = TRUE)
      split(temp, row(temp))} else combn(annotators, 2, simplify = FALSE)
  purrr::map_df(annotator_pairs, ~agreement_pair(annotations, .x)) 
}

# Bubble plot on square matrix of potential annotation values 
crossplot <- function(df, tags_ordered) {
  # WARNING! Hardwired column names "item1" and "item2"
  # WARNING! Hardwired constant "topics_ordered", to be provided from calling environment
  ggplot(df, aes(x = item1, y = item2)) +
    geom_point(aes(size = n, colour = log(n))) + 
    scale_y_discrete(name = NULL, limits = rev(tags_ordered), position = "right") + 
    scale_x_discrete(name = NULL, limits = tags_ordered, position = "top")  + 
    theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, hjust = 0))
}

crossplot_pairwise <- function(annotations, annotators = NULL, margins = TRUE, cutoff = 0, tag_order = c("value", "popularity")) {
  # WARNING! Hardwired column name "n"
  df <- agreement_pairwise(annotations, annotators)
  df <- df[df$n > cutoff,]
  # WARNING! Hardwired column names "annotator1" and "annotator2"
  tag_order <- match.arg(tag_order)
  tags_ordered <- switch(tag_order, 
                         popularity = get_tags_ordered_by_decreasing_popularity(annotations),
                         value = sort(unique(annotations$annotation_value)))
  crossplot(df, tags_ordered) + facet_grid(annotator1 ~ annotator2, margins, switch = "y")
}

# Wrapper to shape the df and then call the plot function
crossplot_from_annotations <- function(annotations, annotators = NULL) {
  df <- agreement(annotations)
  crossplot(df)
}
```

```{r}
topic_occurrences_within_annotator <- 
  topics %>% 
  unite(document_annotator, doc_id, annotator) %>% # pairwise_count will only take a single column for the grouping variable
  widyr::pairwise_count(annotation_value, document_annotator, diag = FALSE)
crossplot(topic_occurrences_within_annotator)
```


## Agreement / disagreement by pairs of annotators

Pairwise coccurrences of topics on the same document, from any annotator, including multiple topic tags applied by same annotator to same document. A given tag will be counted (m-1) times, where m is the number of tags applied by any annotator on a given document, i.e. once for each pair of tags that the tag appears in.

Note that to the extent that there are multiple tags from the same annotator (on a given document) with the tags applied by different annotators, this plot does *not* provide a completely clean measure of agreement between annotators (it conflates cooccurence within annotator).

But I don't see a straightforward how to measure agreement where there are multiple tags per annotator.

```{r}
crossplot_pairwise(topics)
```

Censor cells with very few co-occurrences, to make it visually easier to identify the main features of disagreement.

```{r}
crossplot_pairwise(topics, cutoff = 2)
```


Cooccurrences can also be counted within subsets of annotators. Special cases include within the same annotator, i.e. where annotators apply multiple tags, which do they tend to apply together; and within pairs of annotators.


## Co-occurrences between pairs of annotators.

```{r}
# topic_cooccurrences <- 
#   topics %>% 
#   widyr::pairwise_count(annotation_value, doc_id, diag = TRUE) # loses annotator id

#crossplot(topic_cooccurrences)
crossplot_from_topics(topics)
```


## Document-by-document view

```{r}
topic_unique_annotator <- 
  topics %>% 
  select(doc_id, annotator, annotation_value) %>% 
  group_by(doc_id, annotator) %>% 
  mutate(document_annotator_seq = row_number(),
         annotator_unique = ifelse(document_annotator_seq == 1, annotator, paste0(annotator, "_2"))) %>% 
  ungroup() %>% 
  select(-annotator, -document_annotator_seq) %>% 
  spread(annotator_unique, annotation_value)
topic_unique_annotator
```



## Agreement / disagreement by document

```{r}
topics_per_doc <- 
  topics %>%
  group_by(doc_id) %>% 
  summarise(n_distinct_topic = n_distinct(annotation_value))
```

Complete agreement.

```{r}
topics_per_doc %>% 
  filter(n_distinct_topic == 1)
```

Documents with some disagreement.

```{r}
topic_unique_annotator %>% 
  left_join(topics_per_doc, "doc_id") %>% 
  filter(n_distinct_topic > 1)

```


# Topics

```{r}
### A: impute a non_boE category
plot_marginal_tag_use_by_annotator(topics, tag_order = "popularity")
```

```{r}
crossplot_pairwise(topics, cutoff = 0, tag_order = "popularity")
```




# "Nature" tags

```{r}
natures <- 
  annotations %>% 
  filter(annotation_type == "nature")

plot_marginal_tag_use_by_annotator(natures, tag_order = "popularity")
```

```{r}
crossplot_pairwise(natures, cutoff = 0, tag_order = "popularity")
```


# Sentiment

Extract sentiment tags into a dataframe.

```{r}
sentiments <- 
  annotations %>% 
  filter(annotation_type == "sentiment") %>% 
  select(doc_id, id_annotation, annotator, annotation_value)
```

Map sentiments strings to numbers.

```{r}
# WARNING! Hardwired strings
sentiment_map_chr_to_numeric_correct <- 
  tibble::tribble(
    ~annotation_value, ~score,
    "1. Strongly Negative", -3L,
    "2. Negative"         , -2L,
    "3. Weakly Negative"  , -1L,
    "4. Neutral"          ,  0L,
    "5. Weakly Positive"  , +1L,
    "6. Positive"         , +2L,
    "7. Strongly Positive", +3L,
    "Couldn't score"      , NA_integer_
  )

sentiment_map_chr_to_numeric_20170426 <- 
  tibble::tribble(
    ~annotation_value, ~score,
    "1. Strongly Negative", -3L,
    "2. Negative"         , -2L,
    "3. Weakly Negative"  , -1L,
    "4. Weakly Positive"  , +1L,
    "5. Positive"         , +2L,
    "6. Strongly Positive", +3L,
    "7. Couldn't score"      ,  0L
  )


# Note this will lump unscored documents in with explicit "Couldn't score" tags, so to check for documents lacking a sentiment tag, I would need to do so with NA_character_ early on
sentiments %<>%
  left_join(sentiment_map_chr_to_numeric_20170426, by = "annotation_value") %>% 
  rename(annotation_value_str = annotation_value,
         annotation_value = score) # keep varname generic to ease reuse of other code (would become unnecessary if passing as function argument)
```




## Mean by annotator

Treating sentiment scores as on an interval scale from -3 (strongly negative) to +3 (strongly positive).

```{r}
sentiments %>% 
  group_by(annotator) %>% 
  summarise(mean = mean(as.numeric(annotation_value), na.rm = TRUE))
```

## Association between annotators

```{r}
annotators <- unique(sentiments$annotator)
sentiments %>% 
  select(-annotation_value_str) %>% 
  spread(annotator, annotation_value) %>% 
  select_(.dots = as.list(annotators)) %>%
  #mutate_all(as.integer) %>% 
  cor(method = "spearman",
      use = "pairwise.complete.obs") %>% 
  knitr::kable(digits = 2)
```

```{r}
plot_marginal_tag_use_by_annotator(sentiments, tag_order = "value")
```

```{r}
### Q: How to deal with missing value, as will commonly occur? These should only come up when the topic is not BoE.
crossplot_pairwise(sentiments, cutoff = 0, tag_order = "value")
```


# Level/change tags

```{r}
levels_or_changes <- 
  annotations %>% 
  filter(annotation_type = "level/change")

plot_marginal_tag_use_by_annotator(levels_or_changes)
```



# ARCHIVE

## Completeness

A: adapt for checking whatever our final tagging rules are

Each annotator was supposed to apply one and only one sentiment annotation per document. List any violations.

```{r one_annotation_per_annotator}
### A: modify this so that one and only one tag where there is also a topic tag
sentiments %>% 
  complete(doc_id, annotator) %>% 
  group_by(doc_id, annotator) %>% 
  summarise(n = sum(!is.na(annotation_value))) %>% 
  filter(n != 1) %>% 
  arrange(annotator, doc_id, n)
```

From here on we work with whatever sentiment annotations are available (not assuming a balanced panel).

## Completeness and multiple topic tagging

Distribution of number of topics per annotator-document pair (if any).

```{r}
topics_per_annotator_document <- 
  topics %>% 
  complete(doc_id, annotator) %>% 
  group_by(doc_id, annotator) %>% 
  summarise(n_annotations = sum(!is.na(annotation_value)))
topics_per_annotator_document
```

How often is each annotator applying a given number of topic tags to the same  document?

```{r}
ntopic_per_annotator_document_distribution <- 
  topics_per_annotator_document %>% 
  group_by(annotator, n_annotations) %>% 
  summarise(n = n())
ntopic_per_annotator_document_distribution
```

Annotators who have failed to annotate some documents.

```{r}
ntopic_per_annotator_document_distribution %>% 
  filter(n_annotations == 0) %>% 
  select(annotator)
```

Annotators who have applied multiple tags to some documents.

```{r}
ntopic_per_annotator_document_distribution %>% 
  filter(n_annotations > 1) %>% 
  group_by(annotator) %>%  
  summarise()
```

From here on we work with whatever topic annotations are available (not assuming a balanced panel, or a unique topic per annotator-document pair).
