---
title: "Importing an annotated project from WebAnno"
author: "Peter Eckley"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Importing an annotated project from WebAnno}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
library(webannor)
```

The `webannor` package provides functions to import annotated text files
exported by WebAnno in the UIMA CAS XMI format. The project export format adopted by WebAnno is a nested (hierarchical) file tree.

`webannor' provides functions for various entry-points in the hierarchy. You can import the whole project by starting at the outermost path...

```{r}
get_annotations.WebAnno_project("../data-raw/credibility_example")
```

... or focus on a particular document...

```{r}
get_annotations.WebAnno_document("../data-raw/credibility_example/annotation/s11748.txt")
```

... or a particular annotator of a given document (in case there are multiple annotators per document)...

```{r}
get_annotations.WebAnno_annotator("../data-raw/credibility_example/annotation/s11748.txt/webanno3259911804369659129export.zip")
```

(Alternatively, `get_annotations.WebAnno_XMI()` will directly process the XMI file that exists inside the annotator-specific zip file.)